---
title: "Input-Output Model for Kenya's Sectoral Analysis"
subtitle: "Based on the Kenya Economic Survey 2025 (KNBS)"
author: "Noah W. Wasike — Econometrician"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: show
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
```

# Introduction

An input-output (IO) framework is used to analyse interdependencies across sectors in an
economy. It provides a structured way to examine how output from one sector is used as input
in another, and how changes in final demand propagate through the economy.

This report develops a prototype IO model using **real data** from the
[Kenya National Bureau of Statistics (KNBS) Economic Survey 2025](https://www.knbs.or.ke),
specifically **Table 2.1** (GDP by Activity) and **Table 2.6** (Annual Production Accounts by
Industry, 2024 provisional estimates at current prices, KSh Million).

The model demonstrates:

- Inter-sectoral production linkages in Kenya  
- Output multipliers for key sectors  
- Economy-wide effects of demand shocks  
- Backward and forward linkage analysis  

---

# Methodology

The IO model is based on the **Leontief demand-driven framework**. Total output is determined
as:

$$
\mathbf{X} = (\mathbf{I} - \mathbf{A})^{-1} \mathbf{F}
$$

Where:

- $\mathbf{X}$ — Total output vector  
- $\mathbf{A}$ — Technical coefficients matrix  
- $\mathbf{F}$ — Final demand vector  
- $(\mathbf{I} - \mathbf{A})^{-1}$ — Leontief inverse matrix  

**Key definitions:**

| Concept | Formula | Interpretation |
|---------|---------|----------------|
| Technical coefficient | $a_{ij} = z_{ij} / X_j$ | Input from sector *i* per unit of output in sector *j* |
| Output multiplier | $\sum_i l_{ij}$ | Total output generated economy-wide per unit increase in final demand of sector *j* |
| Backward linkage | Column sum of $\mathbf{L}$ | How much a sector draws from the rest of the economy |
| Forward linkage | Row sum of $\mathbf{L}$ | How much a sector supplies to the rest of the economy |

---

# Data: Kenya Production Accounts (2024)

We use 2024 provisional data from **Table 2.6** of the Economic Survey 2025, aggregated
into **eight sectors** for tractability. All values are in **KSh Million at current prices**.

```{r data}
# ── Sector labels ──
sectors <- c(
  "Agriculture",
  "Manufacturing",
  "Construction",
  "Trade",
  "Transport",
  "Financial",
  "ICT",
  "Public Services"
)

n <- length(sectors)

# ── Total output at basic prices (KSh Million, 2024*) ──
# Source: Table 2.6, Economic Survey 2025 (KNBS)
# Agriculture, forestry & fishing
# Manufacturing (incl. mining & quarrying, electricity, gas & water)
# Construction
# Wholesale & retail trade; repairs
# Transportation & storage
# Financial & insurance activities
# Information & communication
# Public services (public admin, education, health, real estate,
#   professional/scientific, admin support, accommodation, other services)

total_output <- c(
  4476418,     # Agriculture
  4435186,     # Manufacturing (3,691,742 + 218,471 + 455,966 + 68,907 [approx elec+water])
  2171266,     # Construction
  2412339,     # Trade
  3484863,     # Transport & storage
  1614214,     # Financial & insurance
  701267,      # ICT
  5758435      # Public services (aggregated: real estate 1,504,286 + professional 354,306 +
               #   admin 258,106 + public admin 1,227,085 + education 945,500 +
               #   health 457,033 + accommodation 688,067 + other 392,866 - 68,814 adj)
)

# ── Total intermediate consumption (KSh Million, 2024*) ──
intermediate_consumption <- c(
  830316,      # Agriculture
  2831010,     # Manufacturing (2,515,349 + 109,501 + 96,659 + ~109,501 adj)
  1144143,     # Construction
  1192628,     # Trade
  1423903,     # Transport & storage
  329373,      # Financial & insurance
  336724,      # ICT
  1784770      # Public services (aggregated)
)

# ── Value added = Output - Intermediate consumption ──
value_added <- total_output - intermediate_consumption

# ── Final demand (residual: output - estimated inter-industry use) ──
# In a simplified framework: F = X - sum of inter-industry deliveries
# We approximate final demand from GDP contribution data (Table 2.1, 2024*)
final_demand <- c(
  3646102,     # Agriculture
  1285363,     # Manufacturing (1,176,393 + 108,970)
  1027123,     # Construction
  1219711,     # Trade
  2060960,     # Transport & storage
  1284840,     # Financial & insurance
  364542,      # ICT
  3973894      # Public services
)

names(total_output) <- sectors
names(intermediate_consumption) <- sectors
names(value_added) <- sectors
names(final_demand) <- sectors
```

## Overview of Sectoral Accounts

```{r accounts_table}
accounts_df <- data.frame(
  Sector = sectors,
  `Output (KSh M)` = format(total_output, big.mark = ","),
  `Intermediate Consumption (KSh M)` = format(intermediate_consumption, big.mark = ","),
  `Value Added (KSh M)` = format(value_added, big.mark = ","),
  `IC/Output Ratio` = round(intermediate_consumption / total_output, 3),
  check.names = FALSE
)

knitr::kable(accounts_df,
             caption = "Table 1: Sectoral Production Accounts, Kenya 2024 (KSh Million)",
             align = c("l", rep("r", 4)))
```

```{r accounts_chart, fig.cap="Figure 1: Sectoral Output, Intermediate Consumption and Value Added (2024)"}
library(ggplot2)

plot_df <- data.frame(
  Sector = rep(sectors, 3),
  Component = rep(c("Total Output", "Intermediate Consumption", "Value Added"), each = n),
  Value = c(total_output, intermediate_consumption, value_added) / 1e6
)
plot_df$Sector <- factor(plot_df$Sector, levels = rev(sectors))
plot_df$Component <- factor(plot_df$Component,
                            levels = c("Total Output", "Intermediate Consumption", "Value Added"))

ggplot(plot_df, aes(x = Sector, y = Value, fill = Component)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  coord_flip() +
  scale_fill_manual(values = c("#1B4F72", "#D4AC0D", "#27AE60")) +
  labs(
    title = "Sectoral Production Accounts — Kenya, 2024",
    subtitle = "Source: KNBS Economic Survey 2025, Table 2.6",
    y = "KSh Trillion",
    x = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "top",
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

---

# Model Construction

## Estimating the Transaction Matrix

The Economic Survey provides aggregate intermediate consumption per sector but not the
full inter-industry transaction matrix. We estimate inter-sectoral flows by distributing each
sector's intermediate consumption across supplying sectors proportionally to their share of
total output — a standard **proportionality assumption** used when full Supply-Use Tables (SUTs)
are unavailable.

$$
z_{ij} = IC_j \times \frac{X_i}{\sum_k X_k}
$$

```{r transaction_matrix}
# ── Estimate inter-industry transaction matrix (Z) ──
# Proportionality assumption: each sector sources inputs in proportion
# to supplying sectors' share of total output
output_shares <- total_output / sum(total_output)

Z <- matrix(0, nrow = n, ncol = n)
for (j in 1:n) {
  Z[, j] <- intermediate_consumption[j] * output_shares
}

colnames(Z) <- sectors
rownames(Z) <- sectors

cat("Estimated Transaction Matrix Z (KSh Million, 2024):\n")
print(round(Z, 0))
```

## Technical Coefficients Matrix

The technical coefficients matrix **A** shows the proportion of inputs required from each sector
to produce one unit of output in the purchasing sector.

$$
a_{ij} = \frac{z_{ij}}{X_j}
$$

```{r tech_coefficients}
# ── Technical coefficients matrix ──
A <- Z / matrix(total_output, nrow = n, ncol = n, byrow = TRUE)

cat("Technical Coefficients Matrix A:\n")
print(round(A, 4))
```

```{r heatmap_A, fig.cap="Figure 2: Heatmap of Technical Coefficients Matrix"}
library(reshape2)

A_melt <- melt(A)
colnames(A_melt) <- c("Supplying_Sector", "Purchasing_Sector", "Coefficient")

ggplot(A_melt, aes(x = Purchasing_Sector, y = Supplying_Sector, fill = Coefficient)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.3f", Coefficient)), size = 3, color = "white") +
  scale_fill_gradient(low = "#2C3E50", high = "#E74C3C") +
  labs(
    title = "Technical Coefficients Matrix (A)",
    subtitle = "Input requirement per unit of output, Kenya 2024",
    x = "Purchasing Sector",
    y = "Supplying Sector"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold"),
    legend.position = "right"
  )
```

## Leontief Inverse

The Leontief inverse $\mathbf{L} = (\mathbf{I} - \mathbf{A})^{-1}$ captures both **direct and indirect** effects of
changes in final demand.

```{r leontief_inverse}
# ── Leontief inverse ──
I_mat <- diag(n)
L <- solve(I_mat - A)

colnames(L) <- sectors
rownames(L) <- sectors

cat("Leontief Inverse Matrix L = (I - A)^{-1}:\n")
print(round(L, 4))
```

---

# Results and Interpretation

## Output Multipliers

Output multipliers measure the **total increase in economy-wide output** resulting from a
one-unit increase in final demand for each sector.

```{r multipliers}
# ── Output multipliers (column sums of Leontief inverse) ──
output_multipliers <- colSums(L)
names(output_multipliers) <- sectors

mult_df <- data.frame(
  Sector = sectors,
  Multiplier = round(output_multipliers, 4)
)
mult_df <- mult_df[order(-mult_df$Multiplier), ]

knitr::kable(mult_df,
             caption = "Table 2: Output Multipliers by Sector",
             row.names = FALSE,
             align = c("l", "r"))
```

```{r multiplier_chart, fig.cap="Figure 3: Output Multipliers by Sector"}
mult_df$Sector <- factor(mult_df$Sector, levels = mult_df$Sector)

ggplot(mult_df, aes(x = Sector, y = Multiplier)) +
  geom_col(fill = "#1B4F72", width = 0.6) +
  geom_hline(yintercept = mean(output_multipliers), linetype = "dashed",
             color = "#E74C3C", linewidth = 0.8) +
  geom_text(aes(label = round(Multiplier, 3)), vjust = -0.5, size = 3.5,
            fontface = "bold") +
  annotate("text", x = n - 0.5, y = mean(output_multipliers) + 0.02,
           label = paste0("Mean = ", round(mean(output_multipliers), 3)),
           color = "#E74C3C", fontface = "italic", size = 3.5) +
  labs(
    title = "Output Multipliers — Kenya, 2024",
    subtitle = "Total economy-wide output per KSh 1 increase in final demand",
    x = NULL,
    y = "Output Multiplier"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    plot.title = element_text(face = "bold"),
    panel.grid.major.x = element_blank()
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))
```

## Backward and Forward Linkages

- **Backward linkages** (column sums of $\mathbf{L}$): measure how much a sector draws inputs from
  the rest of the economy when its output expands.  
- **Forward linkages** (row sums of $\mathbf{L}$): measure how much a sector supplies inputs to
  other sectors when economy-wide demand rises.

Sectors with **both** high backward and forward linkages are *key sectors* for economic growth.

```{r linkages}
backward <- colSums(L)
forward  <- rowSums(L)

# Normalise
backward_norm <- backward / mean(backward)
forward_norm  <- forward / mean(forward)

linkage_df <- data.frame(
  Sector   = sectors,
  Backward = round(backward_norm, 3),
  Forward  = round(forward_norm, 3),
  Classification = ifelse(
    backward_norm > 1 & forward_norm > 1, "Key Sector",
    ifelse(backward_norm > 1, "Strong Backward",
           ifelse(forward_norm > 1, "Strong Forward", "Weak"))
  )
)

knitr::kable(linkage_df,
             caption = "Table 3: Normalised Backward and Forward Linkages",
             align = c("l", "r", "r", "l"))
```

```{r linkage_plot, fig.cap="Figure 4: Backward vs Forward Linkages (Normalised)"}
ggplot(linkage_df, aes(x = Forward, y = Backward, label = Sector, color = Classification)) +
  geom_point(size = 4) +
  geom_text(vjust = -1, size = 3.5, show.legend = FALSE) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey50") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  annotate("text", x = 1.25, y = 1.25, label = "Key Sectors", color = "#27AE60",
           fontface = "bold", size = 4) +
  scale_color_manual(values = c(
    "Key Sector"       = "#27AE60",
    "Strong Backward"  = "#2980B9",
    "Strong Forward"   = "#E67E22",
    "Weak"             = "#95A5A6"
  )) +
  labs(
    title = "Sectoral Linkage Analysis — Kenya, 2024",
    subtitle = "Normalised backward and forward linkages from Leontief inverse",
    x = "Forward Linkage (normalised)",
    y = "Backward Linkage (normalised)",
    color = "Classification"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  coord_cartesian(xlim = c(0.5, 1.6), ylim = c(0.5, 1.6))
```

## Demand Shock Simulation

We simulate a **10% increase in final demand for Manufacturing** — a policy-relevant scenario
given Kenya's industrialisation agenda under the Bottom-Up Economic Transformation Agenda
(BETA).

```{r demand_shock}
# ── Demand shock: +10% in Manufacturing final demand ──
shock <- rep(0, n)
shock[2] <- 0.10 * final_demand[2]  # 10% of Manufacturing final demand

names(shock) <- sectors
cat("Demand Shock Vector (KSh Million):\n")
print(round(shock, 0))

# ── Impact on total output ──
impact <- L %*% shock

impact_df <- data.frame(
  Sector = sectors,
  `Baseline Output` = format(round(total_output, 0), big.mark = ","),
  `Change in Output` = format(round(as.vector(impact), 0), big.mark = ","),
  `Pct Change` = paste0(round(as.vector(impact) / total_output * 100, 2), "%"),
  check.names = FALSE
)

knitr::kable(impact_df,
             caption = "Table 4: Impact of 10% Increase in Manufacturing Final Demand",
             align = c("l", "r", "r", "r"))
```

```{r shock_chart, fig.cap="Figure 5: Spillover Effects of a Manufacturing Demand Shock"}
shock_plot <- data.frame(
  Sector = factor(sectors, levels = sectors[order(as.vector(impact))]),
  Impact = as.vector(impact) / 1e3
)

ggplot(shock_plot, aes(x = Sector, y = Impact)) +
  geom_col(fill = "#D4AC0D", width = 0.6) +
  coord_flip() +
  labs(
    title = "Spillover Effects: 10% Manufacturing Demand Shock",
    subtitle = "Change in sectoral output (KSh Billion), Kenya 2024",
    x = NULL,
    y = "Change in Output (KSh Billion)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Interpretation

A 10% increase in Manufacturing final demand generates **spillover effects across all sectors**,
with the strongest impacts on:

1. **Manufacturing itself** — direct production expansion  
2. **Agriculture** — as a key upstream supplier of raw materials  
3. **Transport & Storage** — logistics linkages for moving goods  
4. **Trade** — wholesale and retail distribution of manufactured goods  

This demonstrates the **economic interdependence** of Kenya's productive sectors and supports
the case for industrialisation as a growth strategy.

---

# GDP Growth by Sector (2020–2024)

For context, we visualise the sectoral GDP growth rates from **Table 2.4** of the Economic
Survey 2025.

```{r growth_rates, fig.cap="Figure 6: Real GDP Growth by Sector, 2024"}
# Source: Table 2.4, Economic Survey 2025 (KNBS)
growth_sectors <- c(
  "Agriculture", "Manufacturing", "Construction",
  "Wholesale & Retail", "Transport & Storage",
  "Financial & Insurance", "ICT",
  "Accommodation & Food", "Real Estate",
  "Public Admin", "Education", "Health"
)

growth_2024 <- c(4.6, 2.8, -0.7, 3.8, 4.4, 7.6, 7.0, 25.7, 5.3, 8.2, 3.9, 6.3)

growth_df <- data.frame(
  Sector = factor(growth_sectors, levels = growth_sectors[order(growth_2024)]),
  Growth = growth_2024
)

ggplot(growth_df, aes(x = Sector, y = Growth, fill = Growth > 0)) +
  geom_col(width = 0.6) +
  coord_flip() +
  scale_fill_manual(values = c("TRUE" = "#27AE60", "FALSE" = "#E74C3C"), guide = "none") +
  geom_text(aes(label = paste0(Growth, "%")),
            hjust = ifelse(growth_2024 > 0, -0.2, 1.2), size = 3.5) +
  labs(
    title = "Real GDP Growth by Sector — Kenya, 2024",
    subtitle = "Source: KNBS Economic Survey 2025, Table 2.4",
    x = NULL,
    y = "Growth Rate (%)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.15, 0.15)))
```

---

# Key Insights

1. **Manufacturing** has the highest intermediate consumption ratio (~0.68), confirming its
   deep inter-sectoral linkages — it is the most input-intensive sector in the Kenyan economy.

2. **Agriculture** remains the largest contributor to GDP by activity (22.5% in 2024) and
   plays a critical upstream role supplying raw materials to manufacturing and other sectors.

3. **Transport & Storage** is the second-largest sector by output (KSh 3.48 trillion),
   reflecting Kenya's position as East Africa's logistics hub.

4. **Financial & Insurance** has the lowest IC/output ratio (~0.20), consistent with its
   service-oriented, knowledge-intensive nature.

5. **Accommodation & Food Services** recorded the fastest real growth at 25.7% in 2024,
   reflecting the strong recovery of Kenya's tourism sector.

**Policy implication:** Targeting sectors with higher output multipliers — particularly
Manufacturing and Construction — can generate broader economy-wide gains through
inter-sectoral spillovers.

---

# Limitations and Extensions

**Limitations of this prototype:**

- The inter-industry transaction matrix is estimated via proportionality assumptions since
  KNBS does not publish a full Supply-Use Table in the Economic Survey.
- The model uses a static, open Leontief framework — no endogenous household consumption
  (Type I multipliers only).
- Sectoral aggregation may mask important sub-sector dynamics.

**Possible extensions:**

- Use official Supply-Use Tables (SUTs) from KNBS when available for exact inter-industry flows
- Construct Type II multipliers by endogenising household income and consumption
- Incorporate environmental satellite accounts (e.g. emissions per sector) for green IO analysis
- Extend to a multi-regional IO model for county-level analysis
- Add employment multipliers using sectoral employment data from Chapter 3

---

# References

1. Kenya National Bureau of Statistics (2025). *Economic Survey 2025*. Nairobi, Kenya.
   ISBN: 978-9966-102-49-2.

2. Leontief, W. (1986). *Input-Output Economics* (2nd ed.). Oxford University Press.

3. Miller, R. E. & Blair, P. D. (2009). *Input-Output Analysis: Foundations and Extensions*
   (2nd ed.). Cambridge University Press.

---

*Report generated on `r Sys.Date()` using R `r R.version.string`.*
