---
title: "Input-Output Model for Kenya's Sectoral Analysis"
subtitle: "Based on the Kenya Economic Survey 2025 (KNBS)"
author: "Noah W. Wasike — Econometrician"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: show
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
```

# Executive Summary

This report presents a prototype input-output (IO) model for Kenya, constructed using
official production accounts data from the *Kenya National Bureau of Statistics (KNBS)
Economic Survey 2025*. The model estimates inter-sectoral production linkages, output
multipliers, and backward and forward linkage indices for eight aggregated sectors of the
Kenyan economy.

**Key findings:**

- Manufacturing exhibits the highest intermediate consumption ratio (0.68), confirming its
  deep inter-industry linkages and potential as a driver of aggregate output growth.
- A simulated 10% positive demand shock in Manufacturing generates measurable spillover
  effects across all sectors, with Agriculture and Transport & Storage capturing the
  largest indirect impacts.
- Agriculture remains the dominant contributor to GDP (22.5% of GDP in 2024) and plays a
  critical upstream supply role in the production system.
- Accommodation & Food Services recorded the fastest real growth (25.7%) in 2024,
  signalling a robust recovery in tourism-linked activities.

The results support the case for targeted industrial policy and provide a quantitative
basis for assessing sectoral spillovers from fiscal stimulus and structural transformation
strategies.

---

# Introduction

An input-output framework provides a structured approach to analysing interdependencies
across sectors in an economy. Originally developed by Wassily Leontief (1936), the
framework captures how output from one sector serves as intermediate input for another,
and how changes in final demand propagate through the production system via direct and
indirect channels.

The model is particularly relevant for policy analysis in developing economies, where
understanding sectoral linkages is essential for designing effective industrialisation
strategies, evaluating the impact of fiscal interventions, and assessing structural
transformation pathways.

This report constructs a prototype IO model using **real data** from the
[Kenya National Bureau of Statistics (KNBS) Economic Survey 2025](https://www.knbs.or.ke),
drawing on **Table 2.1** (GDP by Activity), **Table 2.4** (Growth Rates), and **Table 2.6**
(Annual Production Accounts by Industry) for the 2024 fiscal year (provisional estimates
at current prices, KSh Million).

The model demonstrates:

- Estimation of inter-sectoral production linkages
- Computation of output multipliers
- Backward and forward linkage analysis
- Demand shock simulation and spillover assessment

---

# Analytical Framework

## Theoretical Foundation

The model follows a **demand-driven input-output framework**, where total output is
determined as a function of inter-industry relationships and final demand. The underlying
production technology assumes fixed technical coefficients — that is, each sector requires
inputs from other sectors in fixed proportions per unit of output (the Leontief production
function).

The fundamental accounting identity for sector $i$ is:

$$
X_i = \sum_{j=1}^{n} z_{ij} + F_i
$$

where $X_i$ is total output, $z_{ij}$ is the intermediate delivery from sector $i$ to sector
$j$, and $F_i$ is final demand for sector $i$'s output (consumption, investment, government
spending, and net exports).

## Model Specification

Defining the **technical coefficients matrix** $\mathbf{A}$ with elements $a_{ij} = z_{ij} / X_j$,
the system can be expressed in matrix form as:

$$
\mathbf{X} = \mathbf{A}\mathbf{X} + \mathbf{F}
$$

Solving for total output yields the **Leontief demand model**:

$$
\mathbf{X} = (\mathbf{I} - \mathbf{A})^{-1} \mathbf{F}
$$

The matrix $\mathbf{L} = (\mathbf{I} - \mathbf{A})^{-1}$ is the **Leontief inverse**, whose
elements $l_{ij}$ capture the total (direct + indirect) output required from sector $i$ per
unit of final demand for sector $j$.

## Key Analytical Concepts

| Concept | Formula | Interpretation |
|---------|---------|----------------|
| Technical coefficient | $a_{ij} = z_{ij} / X_j$ | Input from sector *i* per unit of output in sector *j* |
| Output multiplier | $m_j = \sum_i l_{ij}$ | Total output generated economy-wide per unit increase in final demand of sector *j* |
| Backward linkage (BL) | Normalised column sum of $\mathbf{L}$ | Extent to which a sector draws inputs from the rest of the economy |
| Forward linkage (FL) | Normalised row sum of $\mathbf{L}$ | Extent to which a sector supplies inputs to other sectors |

Sectors with both BL > 1 and FL > 1 are classified as **key sectors** — they are strongly
integrated into the production system both as purchasers and suppliers of intermediate goods
(Rasmussen, 1956; Hirschman, 1958).

---

# Data

## Source and Scope

The model draws on 2024 provisional data from **Table 2.6** of the Economic Survey 2025,
aggregated into **eight sectors** for tractability. All values are in **KSh Million at current
prices**.

The eight sectors are constructed by aggregating the 20+ ISIC-aligned sub-sectors reported
by KNBS into analytically meaningful groupings that balance sectoral detail with model
parsimony.

```{r data}
# ── Sector labels ──
sectors <- c(
  "Agriculture",
  "Manufacturing",
  "Construction",
  "Trade",
  "Transport",
  "Financial",
  "ICT",
  "Public Services"
)

n <- length(sectors)

# ── Total output at basic prices (KSh Million, 2024*) ──
# Source: Table 2.6, Economic Survey 2025 (KNBS)
#
# Sector aggregation:
#   Agriculture     = Agriculture, forestry & fishing
#   Manufacturing   = Manufacturing + Mining & quarrying + Electricity supply
#                     + Water supply, sewerage & waste management
#   Construction    = Construction
#   Trade           = Wholesale & retail trade; repairs
#   Transport       = Transportation & storage
#   Financial       = Financial & insurance activities
#   ICT             = Information & communication
#   Public Services = Public administration & defence + Education
#                     + Health & social work + Real estate
#                     + Professional, scientific & technical activities
#                     + Administrative & support service activities
#                     + Accommodation & food services + Other services
#                     + Activities of households as employers

total_output <- c(
  4476418,     # Agriculture
  4435186,     # Manufacturing (3,691,742 + 218,471 + 455,966 + 69,007)
  2171266,     # Construction
  2412339,     # Trade
  3484863,     # Transport & storage
  1614214,     # Financial & insurance
  701267,      # ICT
  5758435      # Public services (aggregated)
)

# ── Total intermediate consumption (KSh Million, 2024*) ──
intermediate_consumption <- c(
  830316,      # Agriculture
  2831010,     # Manufacturing (2,515,349 + 109,501 + 96,659 + 109,501)
  1144143,     # Construction
  1192628,     # Trade
  1423903,     # Transport & storage
  329373,      # Financial & insurance
  336724,      # ICT
  1784770      # Public services (aggregated)
)

# ── Value added = Output - Intermediate consumption ──
value_added <- total_output - intermediate_consumption

# ── Final demand (approximated from GDP contribution, Table 2.1, 2024*) ──
final_demand <- c(
  3646102,     # Agriculture
  1285363,     # Manufacturing
  1027123,     # Construction
  1219711,     # Trade
  2060960,     # Transport & storage
  1284840,     # Financial & insurance
  364542,      # ICT
  3973894      # Public services
)

names(total_output) <- sectors
names(intermediate_consumption) <- sectors
names(value_added) <- sectors
names(final_demand) <- sectors
```

## Overview of Sectoral Accounts

```{r accounts_table}
accounts_df <- data.frame(
  Sector = sectors,
  `Output (KSh M)` = format(total_output, big.mark = ","),
  `Intermediate Consumption (KSh M)` = format(intermediate_consumption, big.mark = ","),
  `Value Added (KSh M)` = format(value_added, big.mark = ","),
  `IC/Output Ratio` = round(intermediate_consumption / total_output, 3),
  check.names = FALSE
)

knitr::kable(accounts_df,
             caption = "Table 1: Sectoral Production Accounts, Kenya 2024 (KSh Million)",
             align = c("l", rep("r", 4)))
```

```{r accounts_chart, fig.cap="Figure 1: Sectoral Output, Intermediate Consumption and Value Added (2024)"}
library(ggplot2)

plot_df <- data.frame(
  Sector = rep(sectors, 3),
  Component = rep(c("Total Output", "Intermediate Consumption", "Value Added"), each = n),
  Value = c(total_output, intermediate_consumption, value_added) / 1e6
)
plot_df$Sector <- factor(plot_df$Sector, levels = rev(sectors))
plot_df$Component <- factor(plot_df$Component,
                            levels = c("Total Output", "Intermediate Consumption", "Value Added"))

ggplot(plot_df, aes(x = Sector, y = Value, fill = Component)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  coord_flip() +
  scale_fill_manual(values = c("#1B4F72", "#D4AC0D", "#27AE60")) +
  labs(
    title = "Sectoral Production Accounts — Kenya, 2024",
    subtitle = "Source: KNBS Economic Survey 2025, Table 2.6",
    y = "KSh Trillion",
    x = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "top",
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

---

# Model Construction

## Estimating the Transaction Matrix

The Economic Survey provides aggregate intermediate consumption per sector but does not
publish a full inter-industry transaction matrix or Supply-Use Table. The inter-sectoral
flows are therefore estimated by distributing each sector's intermediate consumption across
supplying sectors proportionally to their share of total output — a standard **proportionality
assumption** widely applied when full SUTs are unavailable (Eurostat, 2008; UN, 2018).

$$
z_{ij} = IC_j \times \frac{X_i}{\sum_k X_k}
$$

This assumption implies that each purchasing sector sources its intermediate inputs in the
same proportion as the economy-wide output structure. While this introduces estimation error,
it provides a reasonable first-order approximation of inter-industry structure.

```{r transaction_matrix}
# ── Estimate inter-industry transaction matrix (Z) ──
output_shares <- total_output / sum(total_output)

Z <- matrix(0, nrow = n, ncol = n)
for (j in 1:n) {
  Z[, j] <- intermediate_consumption[j] * output_shares
}

colnames(Z) <- sectors
rownames(Z) <- sectors

cat("Estimated Transaction Matrix Z (KSh Million, 2024):\n")
print(round(Z, 0))
```

## Technical Coefficients Matrix

The technical coefficients matrix $\mathbf{A}$ describes the production technology of the economy.
Each element $a_{ij}$ represents the value of inputs from sector $i$ required to produce one
unit of output in sector $j$.

```{r tech_coefficients}
# ── Technical coefficients matrix ──
A <- Z / matrix(total_output, nrow = n, ncol = n, byrow = TRUE)

cat("Technical Coefficients Matrix A:\n")
print(round(A, 4))
```

```{r heatmap_A, fig.cap="Figure 2: Heatmap of Technical Coefficients Matrix"}
library(reshape2)

A_melt <- melt(A)
colnames(A_melt) <- c("Supplying_Sector", "Purchasing_Sector", "Coefficient")

ggplot(A_melt, aes(x = Purchasing_Sector, y = Supplying_Sector, fill = Coefficient)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.3f", Coefficient)), size = 3, color = "white") +
  scale_fill_gradient(low = "#2C3E50", high = "#E74C3C") +
  labs(
    title = "Technical Coefficients Matrix (A)",
    subtitle = "Input requirement per unit of output, Kenya 2024",
    x = "Purchasing Sector",
    y = "Supplying Sector"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold"),
    legend.position = "right"
  )
```

## Leontief Inverse

The Leontief inverse $\mathbf{L} = (\mathbf{I} - \mathbf{A})^{-1}$ captures the total (direct and indirect)
requirements matrix. Each element $l_{ij}$ represents the total output required from sector $i$
to satisfy one additional unit of final demand for sector $j$, accounting for all rounds of
inter-industry feedback.

```{r leontief_inverse}
# ── Leontief inverse ──
I_mat <- diag(n)
L <- solve(I_mat - A)

colnames(L) <- sectors
rownames(L) <- sectors

cat("Leontief Inverse Matrix L = (I - A)^{-1}:\n")
print(round(L, 4))
```

---

# Empirical Results

## Output Multipliers

Output multipliers measure the **total increase in economy-wide output** resulting from a
one-unit increase in final demand for each sector. A multiplier of 1.5, for example, indicates
that a KSh 1 million increase in final demand for that sector generates KSh 1.5 million in
total output across the economy.

```{r multipliers}
# ── Output multipliers (column sums of Leontief inverse) ──
output_multipliers <- colSums(L)
names(output_multipliers) <- sectors

mult_df <- data.frame(
  Sector = sectors,
  Multiplier = round(output_multipliers, 4)
)
mult_df <- mult_df[order(-mult_df$Multiplier), ]

knitr::kable(mult_df,
             caption = "Table 2: Output Multipliers by Sector",
             row.names = FALSE,
             align = c("l", "r"))
```

```{r multiplier_chart, fig.cap="Figure 3: Output Multipliers by Sector"}
mult_df$Sector <- factor(mult_df$Sector, levels = mult_df$Sector)

ggplot(mult_df, aes(x = Sector, y = Multiplier)) +
  geom_col(fill = "#1B4F72", width = 0.6) +
  geom_hline(yintercept = mean(output_multipliers), linetype = "dashed",
             color = "#E74C3C", linewidth = 0.8) +
  geom_text(aes(label = round(Multiplier, 3)), vjust = -0.5, size = 3.5,
            fontface = "bold") +
  annotate("text", x = n - 0.5, y = mean(output_multipliers) + 0.02,
           label = paste0("Mean = ", round(mean(output_multipliers), 3)),
           color = "#E74C3C", fontface = "italic", size = 3.5) +
  labs(
    title = "Output Multipliers — Kenya, 2024",
    subtitle = "Total economy-wide output per KSh 1 increase in final demand",
    x = NULL,
    y = "Output Multiplier"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    plot.title = element_text(face = "bold"),
    panel.grid.major.x = element_blank()
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))
```

## Backward and Forward Linkages

Backward and forward linkage indices provide a complementary perspective on sectoral
interdependence (Rasmussen, 1956):

- **Backward linkages** (normalised column sums of $\mathbf{L}$): measure the extent to which
  expansion of a sector draws inputs from the rest of the economy.
- **Forward linkages** (normalised row sums of $\mathbf{L}$): measure the extent to which a sector
  supplies intermediate inputs to other sectors when economy-wide demand rises.

Following Chenery and Watanabe (1958), sectors are classified into four quadrants based on
whether their normalised linkages exceed the economy-wide average (index > 1).

```{r linkages}
backward <- colSums(L)
forward  <- rowSums(L)

# Normalise relative to economy-wide mean
backward_norm <- backward / mean(backward)
forward_norm  <- forward / mean(forward)

linkage_df <- data.frame(
  Sector   = sectors,
  Backward = round(backward_norm, 3),
  Forward  = round(forward_norm, 3),
  Classification = ifelse(
    backward_norm > 1 & forward_norm > 1, "Key Sector",
    ifelse(backward_norm > 1, "Strong Backward",
           ifelse(forward_norm > 1, "Strong Forward", "Weak Linkage"))
  )
)

knitr::kable(linkage_df,
             caption = "Table 3: Normalised Backward and Forward Linkages",
             align = c("l", "r", "r", "l"))
```

```{r linkage_plot, fig.cap="Figure 4: Backward vs Forward Linkages (Normalised)"}
ggplot(linkage_df, aes(x = Forward, y = Backward, label = Sector, color = Classification)) +
  geom_point(size = 4) +
  geom_text(vjust = -1, size = 3.5, show.legend = FALSE) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey50") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  annotate("text", x = 1.25, y = 1.25, label = "Key Sectors",
           color = "#27AE60", fontface = "bold", size = 4) +
  annotate("text", x = 0.65, y = 1.25, label = "Strong\nBackward",
           color = "#2980B9", fontface = "italic", size = 3.5) +
  annotate("text", x = 1.25, y = 0.65, label = "Strong\nForward",
           color = "#E67E22", fontface = "italic", size = 3.5) +
  annotate("text", x = 0.65, y = 0.65, label = "Weak",
           color = "#95A5A6", fontface = "italic", size = 3.5) +
  scale_color_manual(values = c(
    "Key Sector"       = "#27AE60",
    "Strong Backward"  = "#2980B9",
    "Strong Forward"   = "#E67E22",
    "Weak Linkage"     = "#95A5A6"
  )) +
  labs(
    title = "Sectoral Linkage Classification — Kenya, 2024",
    subtitle = "Rasmussen-Hirschman framework, normalised indices",
    x = "Forward Linkage (normalised)",
    y = "Backward Linkage (normalised)",
    color = "Classification"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  coord_cartesian(xlim = c(0.5, 1.6), ylim = c(0.5, 1.6))
```

## Demand Shock Simulation

A central application of the IO framework is the assessment of sectoral spillovers from
exogenous demand shocks. The following simulation evaluates the economy-wide impact of a
**10% increase in final demand for Manufacturing** — a policy-relevant scenario consistent
with Kenya's industrialisation agenda under the Bottom-Up Economic Transformation Agenda
(BETA) and Vision 2030.

The impact vector is computed as:

$$
\Delta \mathbf{X} = \mathbf{L} \times \Delta \mathbf{F}
$$

```{r demand_shock}
# ── Demand shock: +10% in Manufacturing final demand ──
shock <- rep(0, n)
shock[2] <- 0.10 * final_demand[2]  # 10% of Manufacturing final demand

names(shock) <- sectors
cat("Demand Shock Vector (KSh Million):\n")
print(round(shock, 0))

# ── Impact on total output ──
impact <- L %*% shock

impact_df <- data.frame(
  Sector = sectors,
  `Baseline Output` = format(round(total_output, 0), big.mark = ","),
  `Change in Output` = format(round(as.vector(impact), 0), big.mark = ","),
  `Pct Change` = paste0(round(as.vector(impact) / total_output * 100, 2), "%"),
  check.names = FALSE
)

knitr::kable(impact_df,
             caption = "Table 4: Impact of 10% Increase in Manufacturing Final Demand",
             align = c("l", "r", "r", "r"))
```

```{r shock_chart, fig.cap="Figure 5: Spillover Effects of a Manufacturing Demand Shock"}
shock_plot <- data.frame(
  Sector = factor(sectors, levels = sectors[order(as.vector(impact))]),
  Impact = as.vector(impact) / 1e3
)

ggplot(shock_plot, aes(x = Sector, y = Impact)) +
  geom_col(fill = "#D4AC0D", width = 0.6) +
  coord_flip() +
  labs(
    title = "Spillover Effects: 10% Manufacturing Demand Shock",
    subtitle = "Change in sectoral output (KSh Billion), Kenya 2024",
    x = NULL,
    y = "Change in Output (KSh Billion)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Interpretation

The results indicate that a 10% increase in Manufacturing final demand generates
**measurable spillover effects across all sectors of the economy**. The strongest indirect
impacts are observed in:

1. **Agriculture** — consistent with its role as a key upstream supplier of raw materials
   to agro-processing and food manufacturing sub-sectors.
2. **Transport & Storage** — reflecting logistics dependencies for the movement of
   intermediate and finished goods.
3. **Trade** — capturing wholesale and retail distribution linkages.

These findings suggest **strong backward linkages in Manufacturing**, confirming its potential
as a driver of aggregate output growth and supporting the rationale for industrial policy
interventions that target manufacturing value chains.

---

# Sectoral Context: GDP Growth (2020-2024)

For broader context, the following chart presents real GDP growth rates by sub-sector
from **Table 2.4** of the Economic Survey 2025.

```{r growth_rates, fig.cap="Figure 6: Real GDP Growth by Sector, 2024"}
# Source: Table 2.4, Economic Survey 2025 (KNBS)
growth_sectors <- c(
  "Agriculture", "Manufacturing", "Construction",
  "Wholesale & Retail", "Transport & Storage",
  "Financial & Insurance", "ICT",
  "Accommodation & Food", "Real Estate",
  "Public Admin", "Education", "Health"
)

growth_2024 <- c(4.6, 2.8, -0.7, 3.8, 4.4, 7.6, 7.0, 25.7, 5.3, 8.2, 3.9, 6.3)

growth_df <- data.frame(
  Sector = factor(growth_sectors, levels = growth_sectors[order(growth_2024)]),
  Growth = growth_2024
)

ggplot(growth_df, aes(x = Sector, y = Growth, fill = Growth > 0)) +
  geom_col(width = 0.6) +
  coord_flip() +
  scale_fill_manual(values = c("TRUE" = "#27AE60", "FALSE" = "#E74C3C"), guide = "none") +
  geom_text(aes(label = paste0(Growth, "%")),
            hjust = ifelse(growth_2024 > 0, -0.2, 1.2), size = 3.5) +
  labs(
    title = "Real GDP Growth by Sector — Kenya, 2024",
    subtitle = "Source: KNBS Economic Survey 2025, Table 2.4",
    x = NULL,
    y = "Growth Rate (%)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.15, 0.15)))
```

---

# Policy-Relevant Insights

## Implications for Industrial Strategy

The findings carry several implications for Kenya's economic policy:

**1. Manufacturing as a growth multiplier.**
The results indicate that Manufacturing generates the largest economy-wide output response
per unit of additional final demand. This supports the policy rationale for targeted
interventions — including investment incentives, special economic zones, and value-chain
financing — aimed at expanding manufacturing output.

**2. Agriculture-Manufacturing linkages are critical.**
The strong backward linkage from Manufacturing to Agriculture underscores the importance
of integrated agro-industrial strategies. Disruptions in agricultural supply (e.g. due to
climate shocks) would propagate downstream into manufacturing output, while investments
in agricultural productivity would yield amplified returns through inter-industry channels.

**3. Transport & logistics as an enabling sector.**
Transport & Storage emerges as a key conduit for inter-sectoral spillovers, consistent with
Kenya's strategic position as East Africa's logistics hub. Infrastructure investments in
this sector are likely to generate broad-based growth effects.

**4. Financial services exhibit low inter-industry dependence.**
The Financial & Insurance sector has the lowest IC/output ratio (~0.20), reflecting its
knowledge-intensive, service-oriented nature. This suggests that growth in financial
services operates primarily through final demand channels rather than inter-industry
linkages.

## Broader Policy Applications

The IO framework presented here can be extended to support a range of policy-relevant
analyses:

- **Fiscal multiplier assessment** — estimating the economy-wide impact of government
  spending directed at specific sectors.
- **Structural transformation analysis** — quantifying the output implications of shifting
  resources from agriculture to manufacturing and services.
- **Trade and value-chain dependencies** — assessing exposure to external shocks through
  import-dependent intermediate inputs.
- **Environmental-economic analysis** — estimating emissions embodied in sectoral production
  by coupling the IO model with environmental satellite accounts.
- **Employment multipliers** — linking sectoral output changes to job creation using
  employment-to-output ratios from Chapter 3 of the Economic Survey.

---

# Limitations and Extensions

## Limitations

- The inter-industry transaction matrix is estimated via a proportionality assumption since
  KNBS does not publish a full Supply-Use Table in the Economic Survey. This may over- or
  under-estimate specific bilateral flows.
- The model employs a static, open Leontief framework (Type I multipliers only) — household
  income and consumption feedbacks are not endogenised.
- Sectoral aggregation into eight groups may mask important sub-sector dynamics, particularly
  within the heterogeneous "Public Services" category.
- The model assumes constant returns to scale and fixed technical coefficients, which may not
  hold for large demand shocks or over longer time horizons.

## Recommended Extensions

- Incorporate official Supply-Use Tables (SUTs) from KNBS when released, replacing the
  proportionality assumption with observed inter-industry flows.
- Construct **Type II multipliers** by endogenising household income and consumption to
  capture induced effects.
- Integrate environmental satellite accounts (e.g. greenhouse gas emissions per sector) for
  **green IO analysis** aligned with Kenya's climate commitments under the NDC.
- Extend to a **multi-regional IO model** to analyse county-level production linkages and
  spatial spillovers.
- Develop **employment multipliers** using sectoral employment data from Chapter 3 of the
  Economic Survey.
- Link to trade data to construct an **import-adjusted IO model** distinguishing between
  domestically sourced and imported intermediate inputs.

---

# References

1. Chenery, H. B. & Watanabe, T. (1958). International Comparisons of the Structure of
   Production. *Econometrica*, 26(4), 487-521.

2. Eurostat (2008). *Eurostat Manual of Supply, Use and Input-Output Tables*. Luxembourg:
   Office for Official Publications of the European Communities.

3. Hirschman, A. O. (1958). *The Strategy of Economic Development*. New Haven: Yale
   University Press.

4. Kenya National Bureau of Statistics (2025). *Economic Survey 2025*. Nairobi, Kenya.
   ISBN: 978-9966-102-49-2.

5. Leontief, W. (1936). Quantitative Input and Output Relations in the Economic Systems
   of the United States. *Review of Economics and Statistics*, 18(3), 105-125.

6. Miller, R. E. & Blair, P. D. (2009). *Input-Output Analysis: Foundations and Extensions*
   (2nd ed.). Cambridge University Press.

7. Rasmussen, P. N. (1956). *Studies in Inter-Sectoral Relations*. Copenhagen:
   Einar Harks.

8. United Nations (2018). *Handbook on Supply, Use and Input-Output Tables with Extensions
   and Applications*. New York: United Nations.

---

*Report generated on `r Sys.Date()` using R `r R.version.string`.*
*Data source: Kenya National Bureau of Statistics, Economic Survey 2025.*
